Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-13 paper/paper.tex
correcting cost model, fixing macros

Formally, the online \emph{Balanced Re-Partitioning} problem (\OBRP) is defined as
follows. We are given a set of $n$ nodes,
initially arbitrarily partitioned into $\ell$~clusters,
each of size~$k$.
We say that a set of nodes $C \subset V$ is \emph{collocated} in a partition $P$
if $P$ assigns them to the same cluster.
\maciek{We can ommit the formal definition: "I.e.,
$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
The input consists of a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
An online algorithm needs to serve the communication request~$(u_t,v_t)$
as soon as it arrives.
Before serving the request,
the online algorithm
can re-partition,
i.e.,
it may move some of the nodes into clusters different than their current assigned clusters.
We assume that
a~communication request between two collocated nodes costs 0,
the cost of a~communication request between two nodes located in different clusters is~1,
and the cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $ALG$,
its cost,
denoted by $ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.
\maciek{Not an expert on commas in english, but here it feels like we have too much. Use some tool to determine (Grammarly?)}

\maciek{Here introduce a restricted model of \PPOBRP, the restricted variant where the input sequence can be perfectly partitioned}

@@

In this section we show a lower bound of $\Omega(k \cdot \ell)$.
The constructed input sequence allows OPT to move to a perfect partition, and hence
the construction also constitues a lower bound for \PPOBRP.
This complements an upper bound of $O(k \cdot \ell)$
for \PPOBRP that we present in Section \ref{sec:upperbound}.

\begin{theorem}
  The competitive ratio of any deterministic algorithm for \OBRP is in $\Omega(k\cdot \ell)$.
\end{theorem}

\begin{proof}
We construct an instance of the problem with $\ell$ clusters 
$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}
Fix any online algorithm \ALG.
We construct the input sequence of requests for \ALG as follows.
First,
we issue $k-2$ (internal) requests so that \ALG can form a component of $k-1$
nodes on the cluster $S_1$.
Recall that internal requests does not incur any cost for \ALG.
If \ALG at any point splits a component\maciek{define splitting}, then
we continue to issue requests between non-collocated nodes of the component until \ALG collocates the nodes of the component.
If \ALG never collocates them then it is not competitive.

>>>>

Formally, the online \emph{Balanced Re-Partitioning} problem (\OBRP{}) is defined as
follows. We are given a set of $n$ nodes,
initially arbitrarily partitioned into $\ell$~clusters,
each of size~$k$.
We say that nodes in $C \subset V$ are \emph{collocated} in a partition $P$
if $P$ assigns them to the same cluster.
\maciek{We can ommit the formal definition: "I.e.,
$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
The input consists of a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
Before serving the request,
the online algorithm
can re-partition,
i.e.,
it may (optionally) move some of the nodes into clusters different than their current assigned clusters.
Once the re-partition is executed,
we serve
a~communication request between two  nodes at cost~0 if they are collocated,
or at cost~1 if they are located in different clusters.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $ALG$,
its cost,
denoted by $ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.
\maciek{Not an expert on commas in english, but here it feels like we have too much. Use some tool to determine (Grammarly?)}

\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}

@@

In this section we show a lower bound of $\Omega(k \cdot \ell)$.
The constructed input sequence allows OPT to move to a perfect partition, and hence
the construction also constitutes a lower bound for \PPOBRP{}.
This complements an upper bound of $O(k \cdot \ell)$
for \PPOBRP{} that we present in Section \ref{sec:upperbound}.

\begin{theorem}
  The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$.
\end{theorem}

\begin{proof}
We construct an instance of the problem with $\ell$ clusters 
$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
Fix any online algorithm \ALG.
We construct the input sequence of requests for \ALG{} as follows.
First,
we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
nodes on the cluster $S_1$.
Recall that internal requests does not incur any cost for \ALG.
If \ALG{}at any point splits a component\maciek{define splitting}, then
we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
If \ALG{} never collocates them then it is not competitive.
